---
title: "test_CFU_calcs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
library(readxl)
library(dplyr)
library(tidyr)
library(stringr)
library(ggbeeswarm)
library(emmeans)
library(ggplot2)
library(stats)
library(ggpubr)
library(broom)
library(purrr)
```


The only data that needs to be input into this file is the file path to the CFUs

Make the pheno data in two different sheets 
Or read the two sets of data separately
```{r}
file_path <- "../data/D-21_CFUs.xlsx"
```

```{r}
# read in the raw data
CFU_excel_raw <- read_xlsx(file_path, col_names = TRUE, sheet = "counts")

# read in the metadata aka CFU_pheno_data
CFU_metadata <- read_xlsx(file_path, col_names = TRUE, sheet = "metadata")
```

Try to combine lung and spleen
Extract the one row of CFU/dilution that we want to look at

This basically says - choose the dilution/CFU that is the closest to 25 CFUs (because normally we'd count the one that's between 5-50 CFUs for accuracy)
```{r}
# selects all columns that start with dilution
tidy_CFU <- function(df, CFU_column, grouping_columns) {
  
  CFU_column = as.name(CFU_column) # use this because I'm making changes to the column
  grouping_columns <- syms(grouping_columns) # use this for grouping
  
  df %>%
    gather(key = dilution, value = CFUs, matches("dilution")) %>%
    mutate(dilution = str_replace(dilution, "dilution_", ""),
           dilution = as.numeric(dilution),
           CFUs = str_replace(!! sym(CFU_column), "TNTC", "NA"),
           CFUs = as.numeric(!! sym(CFU_column))) %>%
    mutate(CFUs_2 = ifelse(CFUs == 0, NA, CFUs),
         sub_CFUs = abs(25 - CFUs_2)) %>%
    group_by(!!!grouping_columns) %>%
    arrange(!!!grouping_columns, sub_CFUs, dilution) %>% 
    slice(1) %>%
    select(-CFUs_2, -sub_CFUs)
}

CFU_excel_clean <- tidy_CFU(CFU_excel_raw, "CFUs", c("group", "mouse", "organ"))


# add in the dilution facotr, resuspend_volume, and percentage
CFU_all_data <- CFU_excel_clean %>%
  mutate(dilution_factor = CFU_metadata$dilution_factor,
         resuspend_volume_mL = CFU_metadata$resuspend_volume_mL,
         percent = case_when(organ == "Lung" ~ CFU_metadata$lung_percent,
                             TRUE ~ CFU_metadata$spleen_percent))


CFU_all_data_clean <- CFU_all_data %>%
  mutate(dilution = as.numeric(dilution),
         CFUs = as.numeric(CFUs),
         dilution_factor = as.numeric(dilution_factor),
         resuspend_volume_mL = as.numeric(resuspend_volume_mL),
         percent = as.numeric(percent)) 
```

Calculate CFUs
```{r}
# set the constants or can pull out from the metadata
dilution_factor <- 5
resuspend_volume_ml <- 0.5
percent <- 0.3

# calculate the CFUs based on the given data
calculate_cfu <- function(df, dilution_factor, resuspend_volume_ml, percent, dilution_column, CFU_column) {
  
  CFU_column = as.name(CFU_column)
  dilution_column = as.name(dilution_column)
  
  df %>%
    mutate(calculated_CFU = (10*!! rlang::sym(CFU_column)*dilution_factor^ !! rlang::sym(dilution_column))*resuspend_volume_mL,
         whole_CFUs = calculated_CFU/percent,
         log_CFUs = log10(whole_CFUs),
         log_CFUs = as.numeric(str_replace(log_CFUs, "-Inf", "0")))
}

calculate_cfu(CFU_all_data_clean, 5, 0.5, 0.3, "dilution", "CFUs")

```


```{r}
D_minus21_Spleen <- D_minus21_CFUs  %>%
  filter(organ == "Spleen")
OneWayFit <- lm(log_CFUs ~ group, data = D_minus21_Spleen )
anova(OneWayFit)
emmeans(OneWayFit, dunnett ~ group)
```

```{r}
D_minus21_Lung <- D_minus21_CFUs  %>%
  filter(organ == "Lung")
OneWayFit <- lm(log_CFUs ~ group, data = D_minus21_Lung )
anova(OneWayFit)
emmeans(OneWayFit, dunnett ~ group)
```

Plot the data
```{r warning= FALSE}
ggplot(D_minus21_CFUs, aes(x = factor(group, levels = c("PBS", "BCG", "Los Injection", "Prop Injection", "PL Injection", "Los Oral", "Prop Oral","PL Oral")), 
                       y= log_CFUs, color = factor(group, levels = c("PBS", "BCG", "Los Injection", "Prop Injection", "PL Injection", "Los Oral", "Prop Oral","PL Oral")))) +
  geom_point() +
  geom_beeswarm(grouponX = TRUE) +
  facet_wrap(~organ) +
  labs(x = "Group", y = "Log10 CFUs", color = "Group") +
  ggtitle("PL Experiment Lung D-21 BCG CFUs")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
# find the average CFUs per mouse organ
avg_CFU_final <- aggregate(log_CFU_df$log_CFUs, by = list(Group_Mouse = tester$Group_Mouse), FUN = mean) %>%
  rename(avg_log_CFUs = x) %>%
  separate(Group_Mouse, into = c("Group", "Mouse"), sep = "_")
```