---
title: "original_cyto_fe"
output: html_document
---

I need to use this one to confirm that what i'm doing isn't drastically altering the results
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This dataset includes mice that were vaccinated with either PBS (control) or BCG. 12 weeks later these mice were infected with *Mycobacterium tuberculosis*. They were evaluated via CFUs for bacterial burden and flow cytometry for immune populations 30, 60, and 90 days post-infection. A Cytek Aurora was used to collect the flow cytometry data.

Data Manipulations
- removed CD69 and FoxP3 markers because spillover
- removed D30 PBS D mouse because outlier

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, the packages must be loaded.
library(flowCore) - need this package if going to transform the data
```{r warning= FALSE, message = FALSE}
library(tibble)
library(openCyto)
library(data.table)
library(ggcyto)
library(grid)
library(pheatmap)
library(stringr)
library(readxl)
library(gridExtra)
library(ggcorrplot)
library(tidyverse)
library(scales)
library(viridis)
library(superheat)
library(kableExtra)
library(ggpubr)
library(broom)
library(jpeg)



library(mycyto)
```

I have created a few functions to make it easier to analyze the data.

- `clean_FMO_after_gate`: renames the columns; it also removes SSC boundary effects

- `statSmoothFunc`: modified from <https://gist.github.com/kdauria/524eade46135f6348140> to add $r^2$ to correlation plots to correlation plots [Amy, for these modified functions, we should consider giving them different names, so that if we put these in a package and someone else loads this and the package with the original function, there won't be conflicts from two functions from separate packages having the same name. Maybe add `corr` or `r2` to the end of the functoin names?]

- `stat_smooth_func`: modified from <https://gist.github.com/kdauria/524eade46135f6348140> pull out $r^2$ for correlation plots [Amy, For this and the previous function, do you think it would be easier for us to create these modified versions by adding a `geom_text` or `geom_label` line to a ggplot object, instead, rather than changing the geoms for the smooth functions?]

- `as.numeric.factor`: convert the factors for the marker names into numeric for plotting [Amy: Why won't `as.numeric` work for this? That's a base R function that might avoid having to create this new function.]

- `viridis_colors`: vector of the colors used for most plots

- `tidy_flow_item`:  to extract data from an item of a 'flowSet' object

- `tidy_flow_set`: applies tidy flow item to all items in a 'flowSet' object


```{r}

clean_FMO_after_gate <- function(FMO_df) {
  FMO_df %>%
  select(ends_with("-A"), -`FSC-A`, `SSC-A`, filename) %>%
  dplyr::rename(`FoxP3` = "APC-A",
         `CD44` = "APC-Fire 750-A",
        `CD103` =  "APC-R700-A",       
         `CD3` = "Alexa Fluor 532-A",
         `Sca1` = "BB515-A",
         `IL_10` = "BV421-A",
         `CD4` = "BV480-A",
         `CD69` = "BV510-A",
         `CD8` = "BV570-A", 
         `CTLA4` = "BV605-A",
         `CD27` = "BV650-A",
         `CD153` = "BV711-A",
         `KLRG1` = "BV785-A",
         `IL_17` = "PE-A",
         `CD122` = "PE-Cy5-A",
         `IFN` = "PE-Cy7-A", 
         `CD62L` = "PE-Dazzle594-A",
         `TNF` = "Pacific Blue-A", 
         `CD28` = "PE-Cy5.5-A",
         `PD1` = "PerCP-eFluor 710-A")  %>%
   na.omit()%>%
  dplyr::filter(`SSC-A` != max(`SSC-A`))

}


# taken from https://gist.github.com/kdauria/524eade46135f6348140
stat_smooth_func <- function(mapping = NULL, data = NULL,
                        geom = "smooth", position = "identity",
                        ...,
                        method = "auto",
                        formula = y ~ x,
                        se = TRUE,
                        n = 80,
                        span = 0.75,
                        fullrange = FALSE,
                        level = 0.95,
                        method.args = list(),
                        na.rm = FALSE,
                        show.legend = NA,
                        inherit.aes = TRUE,
                        xpos = NULL,
                        ypos = NULL) {
  layer(
    data = data,
    mapping = mapping,
    stat = StatSmoothFunc,
    geom = geom,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(
      method = method,
      formula = formula,
      se = se,
      n = n,
      fullrange = fullrange,
      level = level,
      na.rm = na.rm,
      method.args = method.args,
      span = span,
      xpos = xpos,
      ypos = ypos,
      ...
    )
  )
}

StatSmoothFunc <- ggproto("StatSmooth", Stat,
                      
                      setup_params = function(data, params) {
                        # Figure out what type of smoothing to do: loess for small datasets,
                        # gam with a cubic regression basis for large data
                        # This is based on the size of the _largest_ group.
                        if (identical(params$method, "auto")) {
                          max_group <- max(table(data$group))
                          
                          if (max_group < 1000) {
                            params$method <- "loess"
                          } else {
                            params$method <- "gam"
                            params$formula <- y ~ s(x, bs = "cs")
                          }
                        }
                        if (identical(params$method, "gam")) {
                          params$method <- mgcv::gam
                        }
                        
                        params
                      },
                      
                      compute_group = function(data, scales, method = "auto", formula = y~x,
                                               se = TRUE, n = 80, span = 0.75, fullrange = FALSE,
                                               xseq = NULL, level = 0.95, method.args = list(),
                                               na.rm = FALSE, xpos=NULL, ypos=NULL) {
                        if (length(unique(data$x)) < 2) {
                          # Not enough data to perform fit
                          return(data.frame())
                        }
                        
                        if (is.null(data$weight)) data$weight <- 1
                        
                        if (is.null(xseq)) {
                          if (is.integer(data$x)) {
                            if (fullrange) {
                              xseq <- scales$x$dimension()
                            } else {
                              xseq <- sort(unique(data$x))
                            }
                          } else {
                            if (fullrange) {
                              range <- scales$x$dimension()
                            } else {
                              range <- range(data$x, na.rm = TRUE)
                            }
                            xseq <- seq(range[1], range[2], length.out = n)
                          }
                        }
                        # Special case span because it's the most commonly used model argument
                        if (identical(method, "loess")) {
                          method.args$span <- span
                        }
                        
                        if (is.character(method)) method <- match.fun(method)
                        
                        base.args <- list(quote(formula), data = quote(data), weights = quote(weight))
                        model <- do.call(method, c(base.args, method.args))
                        
                        m = model
                        eq <- substitute(~~italic(r)^2~"="~r2, 
                                         list(a = format(coef(m)[1], digits = 3), 
                                              b = format(coef(m)[2], digits = 3), 
                                              r2 = format(summary(m)$r.squared, digits = 3)))
                        func_string = as.character(as.expression(eq))
                        
                        if(is.null(xpos)) xpos = min(data$x)*0.9
                        if(is.null(ypos)) ypos = max(data$y)*0.9
                        data.frame(x=xpos, y=ypos, label=func_string)
                        
                      },
                      
                      required_aes = c("x", "y")
)

viridis_colors <- c("#3F4788FF", "#56C667FF")


as.numeric.factor <- function(x) {as.numeric(levels(x))[x]}



tidy_flow_item <- function(flowSet_item){
  flowSet_item %>% 
    Biobase::exprs() %>% 
    tibble::as_tibble()
}


tidy_flow_set <- function(flowSet){
  flow_set <- flowCore::fsApply(x = flowSet, 
                                FUN = tidy_flow_item)
  flow_set <- dplyr::bind_rows(flow_set, .id = "filename")
}

```

# Data Cleaning

Read in the initial gating strategy, which gates to live lymphocytes, into a `gatingTemplate` class.

```{r}
# Identify the file with the gating strategy 
ws <- list.files("../Cyto-feature_engineering/data/flow/BCG_v_PBS", 
                 pattern = "AmyGating_intracellular.csv", 
                 full = TRUE)
ws
```

```{r}
# View this template
dtTemplate <- fread(ws)
dtTemplate
```

```{r message = FALSE, warning = FALSE}
# Read in the gating strategy to a 'gatingTemplate' object
initial_gate <- gatingTemplate(ws) 
plot(initial_gate)
```

Gating the sample data

I checked that all D30, D60, and D90 samples have Zombie and will run all the way through the gating when separated by day. Looking at the data, some of them are labeled Zombie NIR-A and Zombie Nir-A which means that they are not recognized as the same. - D90 has uncapitalized. I will need to read them in separately and then rename one of the two so that the cases match

```{r warning=FALSE, fig.height=2.5, fig.width=2.5}

fcsFiles <- list.files("../Cyto-feature_engineering/data/flow/BCG_v_PBS/Tcell/Tcell_final/", 
                       pattern = ".fcs", full = TRUE)

# ncdfFlowset object contains row names with the individual samples and column names with the markers/parameters used in the flow cytometer
ncfs <- read.ncdfFlowSet(fcsFiles, channel_alias = data.frame(alias = c("Zombie Nir-A"), channels = c("Zombie_NIR-A, Zombie NIR-A, Zombie Nir-A"))) 


# apply gating set
gs <- GatingSet(ncfs)

# gate the samples
gt_gating(initial_gate, gs)

# View the gates
plotGate(gs[[1]])



# Pull out the gated data
gated_data <- gs_pop_get_data(gs, "live") %>% 
  as.flowSet() 


# initialize the gated data df
all_gated_df = data.frame()

# convert the gated flow files to dataframes and add a filename column
for (i in 1:length(gated_data)) {
  all_data <- gated_data[[i]] %>%
    exprs() %>%
    data.frame() %>%
    mutate(filename = sampleNames(gated_data[i])) 

  all_gated_df <- rbind(all_gated_df, all_data)
}



# confirm that all of the data is here
unique(all_gated_df$filename)

fig_1a <- ggcyto(gs[[3]], aes(x = `FSC.A`, y = `FSC.H`)) + 
  geom_hex(bins = 50) + 
  geom_gate("singlets") +
  scale_fill_viridis_c() +
  xlab("FSC-A") +
  ylab("FSC-H") +
  ggtitle("Singlets") +
  theme_gray() +
  geom_stats(size = 7) + 
  theme(strip.text = element_blank(),
        axis.text = element_text(size = 24),
        axis.title = element_text(size = 37),
        title = element_text(size = 37)) +
  scale_x_continuous(n.breaks = 3)

fig_1a

#ggsave("../figures/fig_1a.jpeg", fig_1a, dpi = 300)

fig_1b <- ggcyto(gs[[3]], aes(x = `FSC-A`, y = `SSC-A`)) + 
  ggcyto_par_set(limits = list(y = c(0, 3.5e6), x = c(-100, 3.5e6)))+
  geom_hex(bins = 80) + 
  geom_gate("lymph") +
  scale_fill_viridis_c() +
  xlab("FSC-A") +
  ylab("SSC-A") +
  ggtitle("Leukocytes") +
  theme_gray() +
  geom_stats(size = 7)+ 
  theme(strip.text = element_blank(),
        axis.text = element_text(size = 24),
        axis.title = element_text(size = 37),
        title = element_text(size = 37))+
  scale_x_continuous(n.breaks = 3)

fig_1b

#ggsave("../figures/fig_1b.jpeg", fig_1b, dpi = 300)

fig_1c <- ggcyto(gs[[3]], aes(x = `Zombie Nir-A`)) + 
  geom_area(stat = "density", fill = "#440154FF") +
  geom_gate("live")   +
  ggcyto_par_set(limits = list(y = c(0, 5e-5), x = c(-1000, 1e5)))+
  xlab("Zombie NIR-A") +
  ylab("Density") +
  ggtitle("Live") +
  theme_gray() +
  geom_stats(size = 7, adjust = 0.45, position = position_nudge(x = 5000)) +
  labs_cyto("both") + 
  theme(strip.text = element_blank(),
        axis.text = element_text(size = 24),
        axis.title = element_text(size = 37),
        title = element_text(size = 37))+
  scale_x_continuous(n.breaks = 3)

fig_1c

#ggsave("../figures/fig_1c.jpeg", fig_1c, dpi = 300)
```

## Data cleaning of FMOs



Read in the flow cytometry data from the FMO experiments.

```{r}
# Identify the file names of all 20 FCS flow cytometry experiments to 
# read in. 
FMO_fcsFiles <- list.files("../Cyto-feature_engineering/data/flow/BCG_v_PBS/Tcell/FMO_Final", 
                       pattern = ".fcs", full = TRUE)
FMO_fcsFiles
```

```{r}
# Read these files into a 'ncdfFlowSet' object. This will taken a minute
# or so to run. The resulting 'ncdfFlowset' object contains row names with the 
# individual samples and column names with the markers/parameters used in the flow cytometer.
ncfs_FMO <- read.ncdfFlowSet(FMO_fcsFiles, channel_alias = data.frame(alias = c("Zombie Nir-A"), channels = c("Zombie_NIR-A, Zombie NIR-A, Zombie Nir-A"))) 
ncfs_FMO
```

Apply the initial gating to filter the data to only measurements on live lymphocyte cells.

```{r}
# Convert to a 'GatingSet' object, so we can apply the initial gating to this data.
gs_FMO <- GatingSet(ncfs_FMO)
gs_FMO
```

```{r fig.height = 8, fig.width = 4}
# Apply the initial gating to this data, to filter to only measurements
# on live lymphocyte cells. This may take a minute.
gt_gating(initial_gate, gs_FMO)

# You can plot the results of this gating with `plotGate`. For example, to plot
# the gating for the first sample, run:
plotGate(gs_FMO[[1]])
```

## Convert data to "tidy data" format

Now that the initial gating has been applied, to limit the data to measurements oflive, singlet lymphocyte cells, we convert the data to a "tidy data" format, to allow us to work with "tidyverse" tools for further analysis and visualization.

```{r}
# Pull out the data from the 'live' node of the gating set (the last node
# in the initial gating strategy).
FMO_gated_data <- gs_pop_get_data(gs_FMO, "live") %>% 
  as.flowSet() 
```

Apply this function to the 'flowSet' of gated FMO data:

```{r}
FMO_gated_data <- tidy_flow_set(FMO_gated_data)
FMO_gated_data
```


```{r}
# confirm that all of the files were read in correctly
unique(FMO_gated_data$filename)
```


```{r}
# note that here the filename and the column marker names need to match exactly
FMO_gated_data_clean <- FMO_gated_data %>%
  select(ends_with("-A"), -`FSC-A`, `SSC-A`, filename) %>%
  dplyr::rename(`FoxP3` = "APC-A",
         `CD44` = "APC-Fire 750-A",
        `CD103` =  "APC-R700-A",       
         `CD3` = "Alexa Fluor 532-A",
         `Sca1` = "BB515-A",
         `IL_10` = "BV421-A",
         `CD4` = "BV480-A",
         `CD69` = "BV510-A",
         `CD8` = "BV570-A", 
         `CTLA4` = "BV605-A",
         `CD27` = "BV650-A",
         `CD153` = "BV711-A",
         `KLRG1` = "BV785-A",
         `IL_17` = "PE-A",
         `CD122` = "PE-Cy5-A",
         `IFN` = "PE-Cy7-A", 
         `CD62L` = "PE-Dazzle594-A",
         `TNF` = "Pacific Blue-A", 
         `CD28` = "PE-Cy5.5-A",
         `PD1` = "PerCP-eFluor 710-A")  %>%
   na.omit()%>%
  dplyr::filter(`SSC-A` != max(`SSC-A`)) %>%
  mutate(filename = str_replace(filename, "FMOs_", "")) %>%
  mutate(filename = str_replace(filename, "_[:digit:]*_[:digit:]*_Unmixed.fcs", "")) %>%
  mutate(filename = str_replace(filename, "_June_6_Redo_Unmixed.fcs", "")) %>%
mutate(filename = str_replace(filename, "-", "_")) %>%
  mutate(filename = str_replace(filename, "TNF_a", "TNF")) %>%
  mutate(filename = str_replace(filename, "IFN_G", "IFN"))

  
right <- filter_FMO(FMO_gated_data_clean)

add_quantile <- get_99(right)

plot_FMOs(right, add_quantile)
```



## Feature engineer the data

Feature cut the data to get all of the possible populations for each file with the cell count and number of cells. Feature engineering is based on the 99%.

Remove FoxP3
Remove CD69 - spreading from Sca1 makes the marker unusable
Remove D30 PBS D - flow cytometry outlier, both within group, timepoint, and all samples
```{r}

all_gated <- all_gated_df %>%
  dplyr::filter(filename != "D30_Lung_Group 1 D_20190218_153300_Unmixed.fcs") %>%
  mutate(Time = str_extract(filename, "D[0-9]*")) %>%
  mutate(Group = str_extract(filename, "Group\\s[0-9]\\s[A-Z]")) %>%
  unite(filename, c("Time", "Group")) %>%
  mutate(filename= str_replace_all(filename, "\\s", "_")) %>%
  select(ends_with(".A"), -`FSC.A`, filename) %>%
  dplyr::rename(`FoxP3` = "APC.A",
         `CD44` = "APC.Fire.750.A",
        `CD103` =  "APC.R700.A",       
         `CD3` = "Alexa.Fluor.532.A",
         `Sca1` = "BB515.A",
         `IL_10` = "BV421.A",
         `CD4` = "BV480.A",
         `CD69` = "BV510.A",
         `CD8` = "BV570.A", 
         `CTLA4` = "BV605.A",
         `CD27` = "BV650.A",
         `CD153` = "BV711.A",
         `KLRG1` = "BV785.A",
         `IL_17` = "PE.A",
         `CD122` = "PE.Cy5.A",
         `IFN` = "PE.Cy7.A", 
         `CD62L` = "PE.Dazzle594.A",
         `TNF` = "Pacific.Blue.A", 
         `CD28` = "PE.Cy5.5.A",
         `PD1` = "PerCP.eFluor.710.A") %>%
  na.omit() %>%
  dplyr::filter(`SSC.A` != max(`SSC.A`)) %>%
  mutate(CD3 = fe(add_quantile, CD3, "CD3"),
         CD4 = fe(add_quantile, CD4, "CD4"),
         CD8 = fe(add_quantile, CD8, "CD8"),
         CD44 = fe(add_quantile, CD44, "CD44"),
         CD103 = fe(add_quantile, CD103, "CD103"),
         Sca1 = fe(add_quantile, Sca1, "Sca1"),
         IL_17 = fe(add_quantile,IL_17, "IL_17"),
         CD69 = fe(add_quantile,CD69, "CD69"),
         CTLA4 = fe(add_quantile,CTLA4, "CTLA4"),
         CD27 = fe(add_quantile,CD27, "CD27"),
         CD153 = fe(add_quantile,CD153, "CD153"),
         KLRG1 = fe(add_quantile,KLRG1, "KLRG1"),
         IFN = fe(add_quantile,IFN, "IFN"),
         FoxP3 = fe(add_quantile,FoxP3, "FoxP3"),
         CD122 = fe(add_quantile,CD122, "CD122"),
         PD1 = fe(add_quantile,PD1, "PD1"),
         CD62L = fe(add_quantile,CD62L, "CD62L"),
         IL_10 = fe(add_quantile,IL_10, "IL_10"),
         CD28 = fe(add_quantile,CD28, "CD28"),
         TNF = fe(add_quantile,TNF, "TNF")) %>%
  select(-Zombie.Nir.A, -AF.A, -SSC.A, -FoxP3, -CD69) %>%
  mutate_at(vars(-group_cols()), list(~tidyr::replace_na(.,0))) %>%
  group_by(CD3, CD4, CD8, CD44, CD103, Sca1, IL_17, CTLA4, CD27, 
           CD153, KLRG1, IFN,  CD122, PD1, CD62L, IL_10, CD28,
           TNF, filename) %>%
  dplyr::mutate(cell_no = n()) %>%
  unique() %>%
  ungroup() %>%
  dplyr::group_by(filename) %>%
  dplyr::mutate(total_count_by_file = sum(cell_no),
         percentage = (100*cell_no / total_count_by_file)) %>%
  ungroup() 

```

# Visualization
## Initial identification of populations plot

We first want to view all of the different cell phenotypes within the data, and then we can filter the data to see the ones that we're interested in, for example, CD3+ cells that constitute >0.5% of total live leukocytes in a sample.

```{r fig.height = 3, fig.width = 2.5}
# this is the order of the columns that we want for the plot
order_of_markers <- c("CD3", "CD4", "CD8",  "CD44", "CD103", "Sca1", "IL_17","CTLA4",
                      "CD27",  "CD153", "KLRG1", "IFN",  "CD122", "PD1", "CD62L",
                      "IL_10", "CD28","TNF")

# to view all of the possible combinations
total_phenotypes <- filter_for_total_pheno(all_gated, marker_order = order_of_markers)


supp_2 <-heatmap_all_pheno(total_phenotypes)

#ggsave("../figures/supp_2.jpeg", supp_2, dpi = 300)

# gives the total number of populations
nrow(total_phenotypes) 
```

## Filter to smaller subset of populations and plot

Want to add annotations to the phenotype plot

- lineage
- cell
- resident

based on expression of different markers. The phenotype plot is then arranged by lineage, cell and resident cells and split by the lineage (I should try to find a way to do this automatically).

```{r fig.width = 4, fig.height = 6}
# view the specific cell phenotypes we're interested in
sample_populations <- all_gated %>%
  dplyr::filter(CD3 == 1 & percentage > 0.5) %>%
  select(-filename, -percentage, -cell_no, -total_count_by_file) %>%
  group_by(CD3, CD4, CD8, CD44, CD103, Sca1, IL_17, CTLA4, CD27, 
           CD153, KLRG1, IFN,  CD122, PD1, CD62L, IL_10, CD28,
           TNF) %>%
  unique() %>%
  ungroup() %>%
  mutate(population = paste0("Pop", 1:length(CD3)))

# add in the percentage of cells in each specific population after the filter
test <- left_join(sample_populations, all_gated)

# expand to add in the data where a mouse may have 0 cells in a certain population identified after the filter
all_options <- expand(test, population, filename)

# add in the population information
add_pops <- left_join(all_options, sample_populations)

# add populations to the original gated data and replace NA percentages with 0
sample_populations_all_groups <- left_join(add_pops, all_gated) %>%
  select(population, filename, percentage) %>%
  mutate_all(list(~tidyr::replace_na(.,0)))

# these are the colors I want for the pheatmap annotations
my_colors <- list(lineage = c(`Double Negative` = "darkturquoise", 
                              `T Helper` = "plum3",
                              `Cytotoxic T` = "hotpink",
                              `Double Positive` = "orange"),
                  cell = c(`Central Mem` = "darkorchid1", `Effector` = "#F8766D",
                `TSCM` = "#00B0F6", `Unknown` = "yellow"),
                resident = c(`Resident`  = "maroon2", `Non-resident` = "azure"))

# take the filtered data and add annotations for the differnt lineages, cell type, and resident status
phenotype_data <- sample_populations %>%
  mutate_all(~as.numeric.factor(.)) %>%
      mutate(lineage = ifelse(CD4 == 0 & CD8 ==0, "Double Negative",
                          ifelse(CD4 == 1 & CD8 ==0,"T Helper",
                          ifelse(CD4 ==0 & CD8 ==1, "Cytotoxic T", 
                                 "Double Positive")))) %>%
  mutate(cell = ifelse(CD44 == 0 & CD62L == 1 & Sca1 == 1, "TSCM", 
                       ifelse(CD44 ==1 & CD62L==0, "Effector", 
                              ifelse(CD44 == 0 & CD62L == 1 & Sca1 == 0, "Naive",
                                     ifelse(CD44==1 & CD62L ==1, "Central Mem",
                                     "Unknown"))))) %>%
  mutate(resident = ifelse(CD103 == 1, "Resident", "Non-resident")) %>%
  select(lineage, cell, resident, CD3, CD4, CD8, CD44, CD62L, Sca1, CD103,  CTLA4, CD27, CD122, CD153, CD28, PD1, KLRG1, IL_17, IL_10, IFN,  TNF) %>%
  arrange(match(lineage, c("Double Negative", "T Helper", "Cytotoxic T", "Double Positive")), cell, resident) %>%
  mutate(population = c(paste0("Pop", 1: nrow(sample_populations)))) %>%
  column_to_rownames("population") %>%
  rename(`IL-17` = `IL_17`,
         `IL-10` = `IL_10`)

# create an annotated data frame to add to pheatmap
phenotype_annot <- phenotype_data %>%
  select(resident, cell, lineage) %>%
  mutate(population = c(paste0("Pop", 1: nrow(sample_populations)))) %>%
  column_to_rownames("population")

# count the number of rows with each phenotype to create gaps in the plot
DN_length <- sum(phenotype_data$lineage == "Double Negative")
T_helper_length <- sum(phenotype_data$lineage == "T Helper")
Cytotoxic_length <- sum(phenotype_data$lineage == "Cytotoxic T")
DP_length <- sum(phenotype_data$lineage == "Double Positive")

# plot the data with different annotations
fig_3a <- pheatmap(phenotype_data[, 4:21], cluster_rows = FALSE, cluster_cols = FALSE,
           labels_row = paste("Pop", 1:nrow(sample_populations)), 
           annotation_row = phenotype_annot,  gaps_row = c(DN_length, DN_length + T_helper_length, DN_length + T_helper_length + Cytotoxic_length),
           cellwidth = 15, cellheight = 15, angle_col = 45, annotation_colors = my_colors,
         color = viridis_colors, cutree_rows = 2, legend = FALSE)

fig_3a

#ggsave("../figures/fig_3a.jpeg", fig_3a, dpi = 300)
```

