---
title: "Test cyto-fe"
author: "Amy Fox"
date: "6/21/20"
output: html_document
---


This dataset includes mice that were vaccinated with either PBS (control) or BCG. 12 weeks later these mice were infected with *Mycobacterium tuberculosis*. They were evaluated via CFUs for bacterial burden and flow cytometry for immune populations 30, 60, and 90 days post-infection. A Cytek Aurora was used to collect the flow cytometry data.

Data Manipulations
- removed CD69 and FoxP3 markers because spillover
- removed D30 PBS D mouse because outlier

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, the packages must be loaded.
library(flowCore) - need this package if going to transform the data
```{r warning= FALSE, message = FALSE}
library(tibble)
library(openCyto)
library(data.table)
library(ggcyto)
library(grid)
library(pheatmap)
library(stringr)
library(readxl)
library(gridExtra)
library(ggcorrplot)
library(tidyverse)
library(scales)
library(viridis)
library(superheat)
library(kableExtra)
library(ggpubr)
library(broom)
library(jpeg)

library(mycyto) # this is my package
# detach(package:mycyto, unload=TRUE) use this to detach and reload
viridis_colors <- c("#3F4788FF", "#56C667FF")
```

I have created a few functions to make it easier to analyze the data.

- `clean_FMO_after_gate`: renames the columns; it also removes SSC boundary effects





# Data Cleaning

Read in the initial gating strategy, which gates to live lymphocytes, into a `gatingTemplate` class.

```{r}
# Identify the file with the gating strategy 
ws <- list.files("../Cyto-feature_engineering/data/flow/BCG_v_PBS", 
                 pattern = "AmyGating_intracellular.csv", 
                 full = TRUE)
ws
```

```{r}
# View this template
dtTemplate <- fread(ws)
dtTemplate
```

```{r message = FALSE, warning = FALSE}
# Read in the gating strategy to a 'gatingTemplate' object
initial_gate <- gatingTemplate(ws) 
plot(initial_gate)
```

Gating the sample data

I checked that all D30, D60, and D90 samples have Zombie and will run all the way through the gating when separated by day. Looking at the data, some of them are labeled Zombie NIR-A and Zombie Nir-A which means that they are not recognized as the same. - D90 has uncapitalized. I will need to read them in separately and then rename one of the two so that the cases match

```{r warning=FALSE, fig.height=2.5, fig.width=2.5}

fcsFiles <- list.files("../Cyto-feature_engineering/data/flow/BCG_v_PBS/Tcell/Tcell_final/", 
                       pattern = ".fcs", full = TRUE)

# ncdfFlowset object contains row names with the individual samples and column names with the markers/parameters used in the flow cytometer
ncfs <- read.ncdfFlowSet(fcsFiles, channel_alias = data.frame(alias = c("Zombie Nir-A"), channels = c("Zombie_NIR-A, Zombie NIR-A, Zombie Nir-A"))) 

# apply gating set
gs <- GatingSet(ncfs)

# gate the samples
gt_gating(initial_gate, gs)

# View the gates
plotGate(gs[[9]])

# Pull out the gated data
gated_data <- gs_pop_get_data(gs, "live") %>% 
  as.flowSet() 

all_gated_df <- convert_gated_flow(gated_data = gated_data)


# confirm that all of the data is here
unique(all_gated_df$filename)

fig_1a <- ggcyto(gs[[3]], aes(x = `FSC.A`, y = `FSC.H`)) + 
  geom_hex(bins = 50) + 
  geom_gate("singlets") +
  scale_fill_viridis_c() +
  xlab("FSC-A") +
  ylab("FSC-H") +
  ggtitle("Singlets") +
  theme_gray() +
  geom_stats(size = 7) + 
  theme(strip.text = element_blank(),
        axis.text = element_text(size = 24),
        axis.title = element_text(size = 37),
        title = element_text(size = 37)) +
  scale_x_continuous(n.breaks = 3)

fig_1a

#ggsave("../figures/fig_1a.jpeg", fig_1a, dpi = 300)

fig_1b <- ggcyto(gs[[3]], aes(x = `FSC-A`, y = `SSC-A`)) + 
  ggcyto_par_set(limits = list(y = c(0, 3.5e6), x = c(-100, 3.5e6)))+
  geom_hex(bins = 80) + 
  geom_gate("lymph") +
  scale_fill_viridis_c() +
  xlab("FSC-A") +
  ylab("SSC-A") +
  ggtitle("Leukocytes") +
  theme_gray() +
  geom_stats(size = 7)+ 
  theme(strip.text = element_blank(),
        axis.text = element_text(size = 24),
        axis.title = element_text(size = 37),
        title = element_text(size = 37))+
  scale_x_continuous(n.breaks = 3)

fig_1b

#ggsave("../figures/fig_1b.jpeg", fig_1b, dpi = 300)

fig_1c <- ggcyto(gs[[3]], aes(x = `Zombie Nir-A`)) + 
  geom_area(stat = "density", fill = "#440154FF") +
  geom_gate("live")   +
  ggcyto_par_set(limits = list(y = c(0, 5e-5), x = c(-1000, 1e5)))+
  xlab("Zombie NIR-A") +
  ylab("Density") +
  ggtitle("Live") +
  theme_gray() +
  geom_stats(size = 7, adjust = 0.45, position = position_nudge(x = 5000)) +
  labs_cyto("both") + 
  theme(strip.text = element_blank(),
        axis.text = element_text(size = 24),
        axis.title = element_text(size = 37),
        title = element_text(size = 37))+
  scale_x_continuous(n.breaks = 3)

fig_1c

#ggsave("../figures/fig_1c.jpeg", fig_1c, dpi = 300)

```

## Data cleaning of FMOs



Read in the flow cytometry data from the FMO experiments.

```{r}
# Identify the file names of all 20 FCS flow cytometry experiments to 
# read in. 
FMO_fcsFiles <- list.files("../Cyto-feature_engineering/data/flow/BCG_v_PBS/Tcell/FMO_Final", 
                       pattern = ".fcs", full = TRUE)
FMO_fcsFiles
```

```{r}
# Read these files into a 'ncdfFlowSet' object. This will taken a minute
# or so to run. The resulting 'ncdfFlowset' object contains row names with the 
# individual samples and column names with the markers/parameters used in the flow cytometer.
ncfs_FMO <- read.ncdfFlowSet(FMO_fcsFiles, channel_alias = data.frame(alias = c("Zombie Nir-A"), channels = c("Zombie_NIR-A, Zombie NIR-A, Zombie Nir-A"))) 

ncfs_FMO
```

Apply the initial gating to filter the data to only measurements on live lymphocyte cells.

```{r}
# Convert to a 'GatingSet' object, so we can apply the initial gating to this data.
gs_FMO <- GatingSet(ncfs_FMO)
gs_FMO
```

```{r fig.height = 8, fig.width = 4}
# Apply the initial gating to this data, to filter to only measurements
# on live lymphocyte cells. This may take a minute.
gt_gating(initial_gate, gs_FMO)

# You can plot the results of this gating with `plotGate`. For example, to plot
# the gating for the first sample, run:
plotGate(gs_FMO[[1]])
```

## Convert data to "tidy data" format

Now that the initial gating has been applied, to limit the data to measurements oflive, singlet lymphocyte cells, we convert the data to a "tidy data" format, to allow us to work with "tidyverse" tools for further analysis and visualization.

```{r}
# Pull out the data from the 'live' node of the gating set (the last node
# in the initial gating strategy).
FMO_gated_data <- gs_pop_get_data(gs_FMO, "live") %>% 
  as.flowSet() 
```

Apply this function to the 'flowSet' of gated FMO data:

```{r}
FMO_gated_data <- tidy_flow_set(FMO_gated_data)
FMO_gated_data
```


```{r}
# confirm that all of the files were read in correctly
unique(FMO_gated_data$filename)
```

This visualization is to look at the initial gating of the FMOs to ensure that the gates are placed correctly.
```{r fig.height = 8, fig.width = 4}
plotGate(gs_FMO[[3]])
```

# Feature Engineering
## FMOs
Pull out the individual FMOS

```{r}
FMO_gated_data <- gs_pop_get_data(gs_FMO, "live") %>% 
  as.flowSet() %>%
  tidy_flow_set()

colnames(FMO_gated_data)

```

## Gate on FMOs with 99% threshold for + and - populations
Feature engineering FMOs and plot

```{r fig.width=8, fig.height=6}

# note that here the filename and the column marker names need to match exactly
FMO_gated_data_clean <- FMO_gated_data %>%
  select(ends_with("-A"), -`FSC-A`, `SSC-A`, filename) %>%
  dplyr::rename(`FoxP3` = "APC-A",
         `CD44` = "APC-Fire 750-A",
        `CD103` =  "APC-R700-A",       
         `CD3` = "Alexa Fluor 532-A",
         `Sca1` = "BB515-A",
         `IL_10` = "BV421-A",
         `CD4` = "BV480-A",
         `CD69` = "BV510-A",
         `CD8` = "BV570-A", 
         `CTLA4` = "BV605-A",
         `CD27` = "BV650-A",
         `CD153` = "BV711-A",
         `KLRG1` = "BV785-A",
         `IL_17` = "PE-A",
         `CD122` = "PE-Cy5-A",
         `IFN` = "PE-Cy7-A", 
         `CD62L` = "PE-Dazzle594-A",
         `TNF` = "Pacific Blue-A", 
         `CD28` = "PE-Cy5.5-A",
         `PD1` = "PerCP-eFluor 710-A")  %>%
   na.omit()%>%
  dplyr::filter(`SSC-A` != max(`SSC-A`)) %>%
  mutate(filename = str_replace(filename, "FMOs_", "")) %>%
  mutate(filename = str_replace(filename, "_[:digit:]*_[:digit:]*_Unmixed.fcs", "")) %>%
  mutate(filename = str_replace(filename, "_June_6_Redo_Unmixed.fcs", "")) %>%
mutate(filename = str_replace(filename, "-", "_")) %>%
  mutate(filename = str_replace(filename, "TNF_a", "TNF")) %>%
  mutate(filename = str_replace(filename, "IFN_G", "IFN"))

  
right <- filter_FMO(FMO_gated_data_clean)

add_quantile <- get_99(right)

plot_FMOs(right, add_quantile)

#ggsave("../figures/fig_2.jpeg", fig_2, dpi = 300)
```


## Feature engineer the data

Feature cut the data to get all of the possible populations for each file with the cell count and number of cells. Feature engineering is based on the 99%.

Remove FoxP3
Remove CD69 - spreading from Sca1 makes the marker unusable
Remove D30 PBS D - flow cytometry outlier, both within group, timepoint, and all samples
```{r}

all_gated <- all_gated_df %>%
  dplyr::filter(filename != "D30_Lung_Group 1 D_20190218_153300_Unmixed.fcs") %>%
  mutate(Time = str_extract(filename, "D[0-9]*")) %>%
  mutate(Group = str_extract(filename, "Group\\s[0-9]\\s[A-Z]")) %>%
  unite(filename, c("Time", "Group")) %>%
  mutate(filename= str_replace_all(filename, "\\s", "_")) %>%
  select(ends_with(".A"), -`FSC.A`, filename) %>%
  dplyr::rename(`FoxP3` = "APC.A",
         `CD44` = "APC.Fire.750.A",
        `CD103` =  "APC.R700.A",       
         `CD3` = "Alexa.Fluor.532.A",
         `Sca1` = "BB515.A",
         `IL_10` = "BV421.A",
         `CD4` = "BV480.A",
         `CD69` = "BV510.A",
         `CD8` = "BV570.A", 
         `CTLA4` = "BV605.A",
         `CD27` = "BV650.A",
         `CD153` = "BV711.A",
         `KLRG1` = "BV785.A",
         `IL_17` = "PE.A",
         `CD122` = "PE.Cy5.A",
         `IFN` = "PE.Cy7.A", 
         `CD62L` = "PE.Dazzle594.A",
         `TNF` = "Pacific.Blue.A", 
         `CD28` = "PE.Cy5.5.A",
         `PD1` = "PerCP.eFluor.710.A") %>%
  na.omit() %>%
  dplyr::filter(`SSC.A` != max(`SSC.A`)) %>%
  mutate(CD3 = fe(add_quantile, CD3, "CD3"),
         CD4 = fe(add_quantile, CD4, "CD4"),
         CD8 = fe(add_quantile, CD8, "CD8"),
         CD44 = fe(add_quantile, CD44, "CD44"),
         CD103 = fe(add_quantile, CD103, "CD103"),
         Sca1 = fe(add_quantile, Sca1, "Sca1"),
         IL_17 = fe(add_quantile,IL_17, "IL_17"),
         CD69 = fe(add_quantile,CD69, "CD69"),
         CTLA4 = fe(add_quantile,CTLA4, "CTLA4"),
         CD27 = fe(add_quantile,CD27, "CD27"),
         CD153 = fe(add_quantile,CD153, "CD153"),
         KLRG1 = fe(add_quantile,KLRG1, "KLRG1"),
         IFN = fe(add_quantile,IFN, "IFN"),
         FoxP3 = fe(add_quantile,FoxP3, "FoxP3"),
         CD122 = fe(add_quantile,CD122, "CD122"),
         PD1 = fe(add_quantile,PD1, "PD1"),
         CD62L = fe(add_quantile,CD62L, "CD62L"),
         IL_10 = fe(add_quantile,IL_10, "IL_10"),
         CD28 = fe(add_quantile,CD28, "CD28"),
         TNF = fe(add_quantile,TNF, "TNF")) %>%
  select(-Zombie.Nir.A, -AF.A, -SSC.A, -FoxP3, -CD69) %>%
  count_calc(marker_vector = c("CD3", "CD4", "CD8", "CD44", "CD103", "Sca1", "IL_17", "CTLA4",
                               "CD27",  "CD153", "KLRG1", "IFN",  "CD122", "PD1", "CD62L",
                               "IL_10", "CD28","TNF", "filename"))

```




# Visualization
## Initial identification of populations plot

We first want to view all of the different cell phenotypes within the data, and then we can filter the data to see the ones that we're interested in, for example, CD3+ cells that constitute >0.5% of total live leukocytes in a sample.

```{r fig.height = 3, fig.width = 2.5}

# this is the order of the columns that we want for the plot
order_of_markers <- c("CD3", "CD4", "CD8",  "CD44", "CD103", "Sca1", "IL_17","CTLA4",
                      "CD27",  "CD153", "KLRG1", "IFN",  "CD122", "PD1", "CD62L",
                      "IL_10", "CD28","TNF")

# to view all of the possible combinations
total_phenotypes <- filter_for_total_pheno(all_gated, marker_order = order_of_markers)


supp_2 <-heatmap_all_pheno(total_phenotypes)

#ggsave("../figures/supp_2.jpeg", supp_2, dpi = 300)

# gives the total number of populations
nrow(total_phenotypes) 
```

## Filter to smaller subset of populations and plot

Want to add annotations to the phenotype plot

- lineage
- cell
- resident

based on expression of different markers. The phenotype plot is then arranged by lineage, cell and resident cells and split by the lineage (I should try to find a way to do this automatically).

```{r fig.width = 4, fig.height = 6}
# view the specific cell phenotypes we're interested in
sample_populations <- all_gated %>%
  dplyr::filter(CD3 == 1 & percentage > 0.5) %>%
  select(-filename, -percentage, -cell_no, -total_count_by_file) %>%
  group_by_if(is.factor) %>%
  unique() %>%
  ungroup() %>%
  mutate(population = paste0("Pop", 1:length(CD3)))

# add in the percentage of cells in each specific population after the filter
test <- left_join(sample_populations, all_gated)

# expand to add in the data where a mouse may have 0 cells in a certain population identified after the filter
all_options <- expand(test, population, filename)

# add in the population information
add_pops <- left_join(all_options, sample_populations)

# add populations to the original gated data and replace NA percentages with 0
sample_populations_all_groups <- left_join(add_pops, all_gated) %>%
  select(population, filename, percentage) %>%
  mutate_all(list(~tidyr::replace_na(.,0)))

# these are the colors I want for the pheatmap annotations
my_colors <- list(lineage = c(`Double Negative` = "darkturquoise", 
                              `T Helper` = "plum3",
                              `Cytotoxic T` = "hotpink",
                              `Double Positive` = "orange"),
                  cell = c(`Central Mem` = "darkorchid1", `Effector` = "#F8766D",
                `TSCM` = "#00B0F6", `Unknown` = "yellow"),
                resident = c(`Resident`  = "maroon2", `Non-resident` = "azure"))

# take the filtered data and add annotations for the differnt lineages, cell type, and resident status
phenotype_data <- sample_populations %>%
  mutate_all(~as.numeric.factor(.)) %>%
      mutate(lineage = ifelse(CD4 == 0 & CD8 ==0, "Double Negative",
                          ifelse(CD4 == 1 & CD8 ==0,"T Helper",
                          ifelse(CD4 ==0 & CD8 ==1, "Cytotoxic T", 
                                 "Double Positive")))) %>%
  mutate(cell = ifelse(CD44 == 0 & CD62L == 1 & Sca1 == 1, "TSCM", 
                       ifelse(CD44 ==1 & CD62L==0, "Effector", 
                              ifelse(CD44 == 0 & CD62L == 1 & Sca1 == 0, "Naive",
                                     ifelse(CD44==1 & CD62L ==1, "Central Mem",
                                     "Unknown"))))) %>%
  mutate(resident = ifelse(CD103 == 1, "Resident", "Non-resident")) %>%
  select(lineage, cell, resident, CD3, CD4, CD8, CD44, CD62L, Sca1, CD103,  CTLA4, CD27, CD122, CD153, CD28, PD1, KLRG1, IL_17, IL_10, IFN,  TNF) %>%
  arrange(match(lineage, c("Double Negative", "T Helper", "Cytotoxic T", "Double Positive")), cell, resident) %>%
  mutate(population = c(paste0("Pop", 1: nrow(sample_populations)))) %>%
  column_to_rownames("population") %>%
  rename(`IL-17` = `IL_17`,
         `IL-10` = `IL_10`)

# create an annotated data frame to add to pheatmap
phenotype_annot <- phenotype_data %>%
  select(resident, cell, lineage) %>%
  mutate(population = c(paste0("Pop", 1: nrow(sample_populations)))) %>%
  column_to_rownames("population")

# count the number of rows with each phenotype to create gaps in the plot
DN_length <- sum(phenotype_data$lineage == "Double Negative")
T_helper_length <- sum(phenotype_data$lineage == "T Helper")
Cytotoxic_length <- sum(phenotype_data$lineage == "Cytotoxic T")
DP_length <- sum(phenotype_data$lineage == "Double Positive")

# plot the data with different annotations
fig_3a <- pheatmap(phenotype_data[, 4:21], cluster_rows = FALSE, cluster_cols = FALSE,
           labels_row = paste("Pop", 1:nrow(sample_populations)), 
           annotation_row = phenotype_annot,  gaps_row = c(DN_length, DN_length + T_helper_length, DN_length + T_helper_length + Cytotoxic_length),
           cellwidth = 15, cellheight = 15, angle_col = 45, annotation_colors = my_colors,
         color = viridis_colors, cutree_rows = 2, legend = FALSE)

fig_3a

#ggsave("../figures/fig_3a.jpeg", fig_3a, dpi = 300)

colnames(all_gated) 
```

Want to get the percentages added so can plot the percentages with the new 

```{r}
# change the population rownames to column 
phenotype_data <- phenotype_data %>%
  rownames_to_column(var = "population")%>%
  rename(`IL_17` = `IL-17`,
         `IL_10` = `IL-10`)

# convert the marker columns of factor class into numeric
all_gated <- all_gated %>%
  mutate_if(is.factor, ~as.numeric.factor(.)) 

# add the percentages from original data to the populations of filtered data
add_perc <- left_join(phenotype_data, all_gated)

# expand the data so can see which files have 0 cells in a phenotype
all_options <- expand(add_perc, population, filename)

# add the populations back 
add_pops <- left_join(all_options, phenotype_data)

# add the percentages back
sample_populations_all_groups <- left_join(add_pops, all_gated) %>%
  select(population, filename, percentage) %>%
  mutate_all(list(~replace_na(., 0))) 

# confirm that the populations were added correctly
unique(sample_populations_all_groups$population)

```

## Correlation plot
Data Visualization

 - Population correlation Plot

```{r fig.height = 6, fig.width = 6}

# correlation plot without the name "pop"
corr <- sample_populations_all_groups %>%
  mutate(population= str_replace(population, "Pop", "")) %>%
  spread(key = population, value = percentage) %>%
  select(-filename) %>%
  select(c(paste0(1:nrow(phenotype_data))))

corr <- round(cor(corr), 1)

# plot the population correlation plot separated by the lineages
fig_3b <- superheat(corr,  row.title = "Populations", column.title = "Populations",  column.title.size = 8, row.title.size = 8, bottom.label.col = c(rep("darkturquoise", DN_length), rep("plum3", T_helper_length), rep("hotpink", Cytotoxic_length), rep("orange", DP_length)), left.label.col = c(rep("darkturquoise", DN_length), rep("plum3", T_helper_length), rep("hotpink", Cytotoxic_length), rep("orange", DP_length)), bottom.label.size = .05, left.label.size = .05, bottom.label.text.size = 5, left.label.text.size = 5)

fig_3b

#ggsave("../figures/fig_3b.jpeg", fig_3b[[2]], dpi = 300)
```

## Time series plot
Data visualization

- time series
- treated v. untreated

```{r fig.width = 6, fig.height= 4.5}
color_for_facets <- phenotype_data %>%
  select(lineage, population) %>%
  mutate(col = ifelse(lineage == "Double Negative", "darkturquoise", 
                      ifelse(lineage == "T Helper", "plum3",
                             ifelse(lineage == "Cytotoxic T", "hotpink", 
                                    ifelse(lineage == "Double Positive", "orange", "brown")))))


# take the data for filtered populations and rename so that plots are pretty
pops_for_plots_average <- sample_populations_all_groups %>%
  separate(filename, into = c("Timepoint", "Mouse", "Group", "Number"), sep = "_") %>%
  dplyr::group_by(population, Timepoint, Group) %>%
  dplyr::mutate(average_percent = mean(percentage)) %>%
  select(-Number, -percentage, -Mouse) %>%
  ungroup() %>%
  unique() %>%
  dplyr::mutate(Group = str_replace(Group, "1", "PBS"),
         Group = str_replace(Group, "2", "BCG")) %>%
  dplyr::mutate(Timepoint = str_replace(Timepoint, "D30", "30"),
         Timepoint = str_replace(Timepoint, "D60", "60"),
         Timepoint = str_replace(Timepoint, "D90", "90")) %>%
  mutate(Group = str_replace(Group, "PBS", "Control"),
         Group = str_replace(Group, "BCG", "Vaccinated"))

facets_for_timeseries <- left_join(pops_for_plots_average, color_for_facets, by = "population")

# convert populations to factor to put in order for the plot
facets_for_timeseries$population <-  factor(facets_for_timeseries$population, 
                                    levels = paste0("Pop", c(1:length(unique(facets_for_timeseries$population)))))


fig_4 <- ggplot(facets_for_timeseries, aes(x = factor(Timepoint, levels = c("30", "60", "90")),
                                 y = average_percent, group = Group, color = Group)) +
  geom_rect(aes(fill = col),xmin = -Inf,xmax = Inf,
            ymin = -Inf,ymax = Inf, alpha = 0.2) +
  scale_fill_identity() +
  geom_point() +
  geom_line() +
  facet_wrap("population", scales = "free", ncol = 7) +
  xlab("Days Post-Infection") +
  ylab("Average Percent of Total Live Leukocytes") +
  scale_color_manual(values = c("black", "white")) +
  theme_gray() +
    theme(axis.text.x = element_text(angle = 45, size = 13, hjust = 1),
        axis.text.y = element_text(size = 13),
        strip.text.x = element_text(size = 10),
        axis.title = element_text(size = 17),
        title = element_text(size = 20),
        legend.text = element_text(size=18),
        legend.key.size = unit(1.5, "line"))


fig_4
#ggsave("../figures/fig_4.jpeg", fig_4, dpi = 300)
```

## Box plot and Bar plot

Extra visualizations
```{r fig.width = 6.5, fig.height= 6.5}
# Vaccinated v. unvaccinated for Day 30 bar
facets_for_vacc_v_unvacc <- facets_for_timeseries %>%
  filter(Timepoint == 30)
  
# bar plot
supp_3a <- ggplot(facets_for_vacc_v_unvacc, aes(x = Group,
                                 y = average_percent, group = Group, fill = Group)) +
  geom_rect(aes(fill = c(col)),xmin = -Inf,xmax = Inf,
            ymin = -Inf,ymax = Inf, alpha = .6) +
  geom_bar(stat = "identity") +
  facet_wrap("population", scales = "free", ncol = 7) +
  scale_fill_manual(values = c("Control" = "black", "Vaccinated" = "white", "darkturquoise" = "darkturquoise",  "hotpink" = "hotpink","orange" = "orange","plum3" = "plum3")) +
  xlab("Vaccination Status") +
  ylab("Average Percent of Cells") +
  ggtitle("Lung Cell Percentage of Total Live Leukocytes") +
  theme_gray() +
    theme(axis.text.x = element_text(angle = 40, size = 15, hjust = 1),
        axis.text.y = element_text(size = 15),
        strip.text.x = element_text(size = 13),
        axis.title = element_text(size = 20),
        title = element_text(size = 20),
        legend.text = element_text(size=18),
        legend.key.size = unit(1.5, "line"),
        legend.position = "none")
supp_3a
#ggsave("../figures/supp_3a.jpeg", supp_3a, dpi = 300)

# boxplot
facets_for_boxplot <- sample_populations_all_groups %>%
  separate(filename, into = c("Timepoint", "Mouse", "Group", "Number"), sep = "_") %>%
  dplyr::group_by(population, Timepoint, Group) %>%
  dplyr::mutate(average_percent = mean(percentage)) %>%
  select(-Number, -Mouse) %>%
  ungroup() %>%
  unique() %>%
  dplyr::mutate(Group = str_replace(Group, "1", "PBS"),
         Group = str_replace(Group, "2", "BCG")) %>%
  dplyr::mutate(Timepoint = str_replace(Timepoint, "D30", "30"),
         Timepoint = str_replace(Timepoint, "D60", "60"),
         Timepoint = str_replace(Timepoint, "D90", "90")) %>%
  filter(Timepoint == 30) %>%
  mutate(Group = str_replace(Group, "PBS", "Control"),
         Group = str_replace(Group, "BCG", "Vaccinated"))

# add background colors
facets_for_boxplot <- left_join(facets_for_boxplot, color_for_facets, by = "population")

# convert populations to factor to put in order for the plot
facets_for_boxplot$population <-  factor(facets_for_boxplot$population, 
                                    levels = paste0("Pop", c(1:length(unique(facets_for_boxplot$population)))))

# boxplot
supp_3b <- ggplot(facets_for_boxplot, aes(x = Group,
                                 y = percentage, group = Group, fill = Group)) +
  geom_rect(aes(fill = c(col)),xmin = -Inf,xmax = Inf,
            ymin = -Inf,ymax = Inf, alpha = .2) +
  geom_boxplot() +
  facet_wrap("population", scales = "free", ncol = 7) +
  scale_fill_manual(values = c("Control" = "black", "Vaccinated" = "white", "darkturquoise" = "darkturquoise",  "hotpink" = "hotpink","orange" = "orange","plum3" = "plum3")) +
  xlab("Vaccination Status") +
  ylab("Percent of Cells") +
  ggtitle("Lung Cell Percentage of Total Live Leukocytes") +
  theme_gray() +
    theme(axis.text.x = element_text(angle = 40, size = 15, hjust = 1),
        axis.text.y = element_text(size = 15),
        strip.text.x = element_text(size = 13),
        axis.title = element_text(size = 20),
        title = element_text(size = 20),
        legend.text = element_text(size=18),
        legend.key.size = unit(1.5, "line"),
        legend.position = "none") 

supp_3b
#ggsave("../figures/supp_3b.jpeg", supp_3b, dpi = 300)

# violin plot
ggplot(facets_for_boxplot, aes(x = Group,
                                 y = percentage, group = Group, fill = Group)) +
  geom_rect(aes(fill = c(col)),xmin = -Inf,xmax = Inf,
            ymin = -Inf,ymax = Inf, alpha = .2) +
  geom_violin() +
  facet_wrap("population", scales = "free", ncol = 7) +
  scale_fill_manual(values = c("Control" = "black", "Vaccinated" = "white", "darkturquoise" = "darkturquoise",  "hotpink" = "hotpink","orange" = "orange","plum3" = "plum3")) +
  xlab("Vaccination Status") +
  ylab("Percent of Cells") +
  ggtitle("Lung Cell Percentage of Total Live Leukocytes") +
  theme_gray() +
    theme(axis.text.x = element_text(angle = 40, size = 13, hjust = 1),
        axis.text.y = element_text(size = 13),
        strip.text.x = element_text(size = 10),
        axis.title = element_text(size = 20),
        title = element_text(size = 20),
        legend.text = element_text(size=18),
        legend.key.size = unit(1.5, "line"),
        legend.position = "none")
```

# Hypothesis testing
## Linear regression
## Multiple comparison testing: Benjamini and Hochberg
Data Visualization

- CFU Correlation

```{r fig.width = 4.5, fig.height= 5}

# read in the CFU data
CFUs <- read_xlsx("../data/M.smeg_CFU.xlsx") %>%
  dplyr::filter(Organ == "Lung") %>%
  dplyr::filter(Group == "Group1" | Group == "Group2")

# clean the flow data and prepare to join with CFU
pops_for_CFUs <- sample_populations_all_groups %>%
  separate(filename, into = c("Timepoint", "Mouse", "Group", "Number"), sep = "_") %>%
  dplyr::mutate(Timepoint = str_replace(Timepoint, "D", "")) %>%
  dplyr::mutate(Timepoint = as.numeric(Timepoint)) %>%
  unite(Group, c(Mouse, Group)) %>%
  dplyr::mutate(Group = str_replace(Group, "_", "")) 

# join together the CFU data and the population data
pops_CFUs <- inner_join(pops_for_CFUs, CFUs, by = c("Group", "Number", "Timepoint")) 

facets_for_CFUs <- left_join(pops_CFUs, color_for_facets, by = "population")

facets_for_CFUs$population <-  factor(facets_for_CFUs$population, 
                                    levels = paste0("Pop", c(1:length(unique(facets_for_CFUs$population)))))

ggplot(facets_for_CFUs, aes(percentage, CFU)) +
  geom_rect(aes(fill = col),xmin = -Inf,xmax = Inf,
            ymin = -Inf,ymax = Inf, alpha = 0.05) + 
  scale_fill_identity() +
  geom_point(color = "black") +
  geom_smooth(method = "lm", se = FALSE, color = "#3F4788FF") +
  facet_wrap(~population, scales = "free_x", ncol = 6) +
  stat_smooth_func(geom= "text", method = "lm", hjust = 0, parse = TRUE, color = "white") +
  xlab("Percentage of Cells") +
  ylab("log10 CFU") +
  ggtitle("Population and CFU Linear regression") +
  theme(axis.text.x = element_text(size = 9, hjust = 1, angle = 45),
        axis.text.y = element_text(size = 9),
        axis.title = element_text(size = 15),
        title = element_text(size = 17))

#ggsave("../figures/fig_5b.jpeg", dpi = 300)

fitted_models <- facets_for_CFUs %>% 
  group_by(population) %>% 
  do(model = lm(percentage ~ CFU, data = .)) %>%
  tidy(model) %>%
  filter(term == "CFU") %>%
  select(population, p.value) %>%
  rename(p.value = p.value) %>%
  ungroup()

fitted_models %>%
  mutate(p.val.adj = p.adjust(fitted_models$p.value, "BH")) %>%
  mutate("Significance" = p.val.adj < 0.05) %>%
  rename("p-value" = p.value,
         "Adjusted p-value" = p.val.adj) %>%
  kable(align = "c") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  row_spec(0:nrow(fitted_models), color = "black")

# me testing out

fitted_models <- facets_for_CFUs %>% 
  group_by(population) %>% 
  do(model = lm(percentage ~ CFU, data = .)) %>%
  glance(model) %>%
  select(population, r.squared, p.value) %>%
  ungroup()

fitted_models %>%
  mutate(p.val.adj = p.adjust(fitted_models$p.value, "BH")) %>%
  mutate("Significance" = p.val.adj < 0.05) %>%
  rename("p-value" = p.value,
         "r squared" = r.squared,
         "Adjusted p-value" = p.val.adj) %>%
  kable(align = "c") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 15) %>%
  row_spec(0:nrow(fitted_models), color = "black")

```


Visualization for individual populations
```{r}
pop_pick <- facets_for_CFUs %>%
  filter(population == "Pop32")

ggplot(pop_pick, aes(percentage, CFU)) +
  geom_rect(aes(fill = col),xmin = -Inf,xmax = Inf,
            ymin = -Inf,ymax = Inf, alpha = 0.05) + 
  scale_fill_identity() +
  geom_point(color = "black", size = 4) +
  geom_smooth(method = "lm", se = FALSE, color = "#3F4788FF") +
  stat_smooth_func(geom= "text", method = "lm", hjust = 0, parse = TRUE, color = "white", size = 8) +
  xlab("Percentage of Cells") +
  ylab("log10 CFU") +
  ggtitle("Population 32 and CFU Linear regression") +
  theme(axis.text.x = element_text(size = 13, hjust = 1),
        axis.text.y = element_text(size = 13),
        axis.title = element_text(size = 15),
        title = element_text(size = 17))

pop_timeseries <- facets_for_timeseries  %>%
  filter(population == "Pop3")

ggplot(pop_timeseries, aes(x = factor(Timepoint, levels = c("30", "60", "90")),
                                 y = average_percent, group = Group, color = Group)) +
  geom_rect(aes(fill = col),xmin = -Inf,xmax = Inf,
            ymin = -Inf,ymax = Inf, alpha = 0.2) +
  scale_fill_identity() +
  geom_point(size = 4) +
  geom_line(size = 2) +
  xlab("Days Post-Infection") +
  ylab("Average Percent of Total Live Leukocytes") +
  ggtitle("Pop 3") +
  scale_color_manual(values = c("black", "white")) +
  theme_gray() +
    theme(axis.text.x = element_text(angle = 45, size = 13, hjust = 1),
        axis.text.y = element_text(size = 13),
        strip.text.x = element_text(size = 10),
        axis.title = element_text(size = 17),
        title = element_text(size = 20),
        legend.text = element_text(size=18),
        legend.key.size = unit(1.5, "line"))

```

Different timepoints for hypothesis
```{r}

# D30
facets_for_CFUs_D30 <- facets_for_CFUs %>%
  filter(Timepoint == 30)

ggplot(facets_for_CFUs_D30, aes(percentage, CFU)) +
  geom_rect(aes(fill = col),xmin = -Inf,xmax = Inf,
            ymin = -Inf,ymax = Inf, alpha = 0.2) + 
  scale_fill_identity() +
  geom_point(color = "black") +
  geom_smooth(method = "lm", se = FALSE, color = "#3F4788FF") +
  facet_wrap(~population, scales = "free_x", ncol = 9) +
  stat_smooth_func(geom= "text", method = "lm", hjust = 0, parse = TRUE) +
  scale_x_continuous(breaks = scales::pretty_breaks(3)) +
  xlab("Percentage of Cells") +
  ylab("log10 CFU") +
  ggtitle("Day 30: Population and CFU Linear regression") +
  theme(axis.text.x = element_text(size = 9, hjust = 1),
        axis.text.y = element_text(size = 9),
        axis.title = element_text(size = 15),
        title = element_text(size = 17))

# D60
facets_for_CFUs_D60 <- facets_for_CFUs %>%
  filter(Timepoint == 60)

ggplot(facets_for_CFUs_D60, aes(percentage, CFU)) +
  geom_rect(aes(fill = col),xmin = -Inf,xmax = Inf,
            ymin = -Inf,ymax = Inf, alpha = 0.2) + 
  scale_fill_identity() +
  geom_point(color = "black") +
  geom_smooth(method = "lm", se = FALSE, color = "#3F4788FF") +
  facet_wrap(~population, scales = "free_x", ncol = 9) +
  stat_smooth_func(geom= "text", method = "lm", hjust = 0, parse = TRUE) +
  scale_x_continuous(breaks = scales::pretty_breaks(3)) +
  xlab("Percentage of Cells") +
  ylab("log10 CFU") +
  ggtitle("Day 60: Population and CFU Linear regression") +
  theme(axis.text.x = element_text(size = 9, hjust = 1),
        axis.text.y = element_text(size = 9),
        axis.title = element_text(size = 15),
        title = element_text(size = 17))

# D90
facets_for_CFUs_D90 <- facets_for_CFUs %>%
  filter(Timepoint == 90)

ggplot(facets_for_CFUs_D90, aes(percentage, CFU)) +
  geom_rect(aes(fill = col),xmin = -Inf,xmax = Inf,
            ymin = -Inf,ymax = Inf, alpha = 0.2) + 
  scale_fill_identity() +
  geom_point(color = "black") +
  geom_smooth(method = "lm", se = FALSE, color = "#3F4788FF") +
  facet_wrap(~population, scales = "free_x", ncol = 9) +
  stat_smooth_func(geom= "text", method = "lm", hjust = 0, parse = TRUE) +
  scale_x_continuous(breaks = scales::pretty_breaks(3)) +
  xlab("Percentage of Cells") +
  ylab("log10 CFU") +
  ggtitle("Day 90: Population and CFU Linear regression") +
  theme(axis.text.x = element_text(size = 9, hjust = 1),
        axis.text.y = element_text(size = 9),
        axis.title = element_text(size = 15),
        title = element_text(size = 17))
```


Confirm pipeline populations with manual gating

```{r dpi = 300}

# read in and clean manual gating percentage data
manual_gating_pop3 <- read_xlsx("../data/flow/BCG_v_PBS/manual_gating_pop3.xlsx") %>%
  dplyr::rename(`Sample` = "Sample:",
         live_leuk = "Singlets/Leukocytes/Live | Count",
         pop3 = "Singlets/Leukocytes/Live/CD3+/CD44+/Sca1+/IFN+/CD4-/CD8-/CD27-/CD28-/CD62L-/CD103-/CD122-/CD153-/CTLA4-/IL10-/IL17-/KLRG1-/PD1-/TNF- | Count",
         percent_manual = "Percentage") %>%
  filter(Sample != "D30_Lung_Group 1 D_20190218_153300_Unmixed.fcs") %>%
  mutate(Sample = str_replace_all(Sample, " ", "_")) %>%
  separate(Sample, into = c("Timepoint", "Organ", "group", "Group", "Number", "Date", "random", "Unmix"), sep = "_") %>%
  mutate(Timepoint = str_replace_all(Timepoint, "D", "")) %>%
  select(Timepoint, Group, Number, percent_manual) %>%
  mutate(population = rep("Pop3", 24))

# prepare the flow CFU data for joining with manual gating
pipeline_pop3 <- pops_CFUs %>%
  dplyr::filter(population == "Pop3") %>%
  mutate(Group = str_replace_all(Group, "Group", "")) %>%
  mutate(Timepoint = as.character(Timepoint)) %>%
  select(-Organ, -CFU) %>%
  dplyr::rename(percent_pipeline = percentage)

# should add absolute value of the percent_pipeline-percent_manual for error (abs)
# join the manual and pipeline percentage data
comparison <- full_join(manual_gating_pop3, pipeline_pop3, by = c("Timepoint", "Group", "Number", "population")) %>%
  mutate(difference = round(percent_pipeline - percent_manual, digits = 2)) %>%
  mutate(error = round(100*(percent_pipeline - percent_manual)/percent_manual, digits = 2)) %>%
  mutate(Group = str_replace_all(Group, "1", "PBS")) %>%
  mutate(Group = str_replace_all(Group, "2", "BCG")) %>%
  mutate(percent_manual = round(percent_manual, digits = 2),
         percent_pipeline = round(percent_pipeline, digits = 2)) %>%
  mutate(absolute_difference = abs(difference),
         average_difference = mean(absolute_difference)) %>%
    select(population, Timepoint, Group, Number, percent_manual, percent_pipeline, absolute_difference, average_difference) 

# print out the comparison data in a pretty table
kable1 <- comparison %>%
  mutate(`% Based On Manual Gating` = percent_manual,
         `% Based on Automated Pipeline` = percent_pipeline,
         `Absolute Difference` = absolute_difference,
         `Average Difference` = average_difference) %>%
  mutate(Group = str_replace(Group, "PBS", "Control"),
         Group= str_replace(Group, "BCG", "Vaccinated")) %>%
  select(Timepoint, Group, Number, `% Based On Manual Gating`, `% Based on Automated Pipeline`, `Absolute Difference`, `Average Difference`) %>%
  kable(align = "c") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 25) %>%
  row_spec(0:nrow(comparison), color = "black")

kable1
#save_kable(kable1, file = "../figures/supp_4b.jpeg")


```

```{r}

# compare the two gating types with manual v. pipeline plot
supp_4c <- ggplot(comparison, aes(x = percent_pipeline, y = percent_manual)) +
  geom_point(size = 3) +
  geom_abline(slope = 1, intercept = 0, size = 1) +
  stat_cor(aes(x = percent_pipeline, y = percent_manual), data = comparison, "spearman", size = 7) +
  ggtitle("Population 3: Manual vs. Pipeline Percentage") +
  xlab("Pipeline Percent") +
  ylab("Manual Percent") +
  theme_classic()+
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 15),
        title = element_text(size = 15))
supp_4c

#ggsave("../figures/supp_4c.jpeg", supp_4c, dpi = 300)
```

```{r}

my_colors <- list(lineage = c(`Double Negative` = "darkturquoise", 
                              `T Helper` = "plum3",
                              `Cytotoxic T` = "hotpink",
                              `Double Positive` = "orange"),
                  cell = c(`Central Mem` = "darkorchid1", `Effector` = "#F8766D",
                `TSCM` = "#00B0F6", `Unknown` = "yellow"),
                resident = c(`Resident`  = "maroon2", `Non-resident` = "azure"))


plot(NULL ,xaxt='n',yaxt='n',bty='n',ylab='',xlab='', xlim=0:1, ylim=0:1)
legend("topleft", legend =c("Negative", "Positive"), pch= 22, pt.cex=4.5, cex=1.5, bty='n',
    col = "darkgray", pt.bg = viridis_colors)
mtext("expression", at=0.1, cex=1.6, font = 2)

plot(NULL ,xaxt='n',yaxt='n',bty='n',ylab='',xlab='', xlim=0:1, ylim=0:1)
legend("topleft", legend =c("Double Negative", "T Helper", "Cytotoxic T", "Double Positive"), 
       pch= 22, pt.cex=4.5, cex=1.5, bty='n',
    col = "darkgray", pt.bg = c("darkturquoise", "plum3", "hotpink", "orange"))
mtext("lineage", at=0.05, cex=1.6, font = 2)

```



```{r}
sessionInfo()
```


